---
title: "Chapter4"
author: "Pinja Mero"
date: "27/11/2021"
output: html_document
---



## Chapter 4: Clustering and classification

### Analysis exercise

#### 2.
 Checking if I allready have a MASS package:
```{r}
a<-installed.packages()
packages<-a[,1] 
is.element("MASS", packages)
```
 
MASS Package exists, load the library and observe the dataset:
```{r}
library(MASS)
str(Boston)
d <- dim(Boston)
d
```

The dataset contains 14 variables and 506 observations. It describes the socio-economic conditions in Boston by housing values of suburbs.


#### 3.


```{r}
library(corrplot)
cormat <- cor(Boston);
corrplot(cormat, method = 'circle')
```
Some correlation can be observed between the variables.

Summaries:
```{r}
B_sum <- setNames(data.frame(matrix(ncol = 6, nrow = 0)), 
              c('Min', '1st Qu', 'Median', 'Mean', '3rd Qu', 'Max'))     #set the empty data frame
for(i in 1:d[2]) {                              # For-loop to create the 14 summary rows
  new <- round(summary(Boston[,i]), digits = 2)   # Create new summary row
  B_sum[nrow(B_sum) + 1, ] <- new                         # Append new row
}
row.names(B_sum) <- colnames(Boston)                  # Set row names
B_sum
```

#### 4.
Summary of the variables (in data frame to look a bit nicer)
```{r}
B_scaled <- data.frame(scale(Boston))

Bsc_sum <- setNames(data.frame(matrix(ncol = 6, nrow = 0)), 
              c('Min', '1st Qu', 'Median', 'Mean', '3rd Qu', 'Max'))     #set the empty data frame
for(i in 1:d[2]) {                              # For-loop to create the 14 summary rows
  new <- round(summary(B_scaled[,i]), digits = 2)   # Create new summary row
  Bsc_sum[nrow(Bsc_sum) + 1, ] <- new                         # Append new row
}
row.names(Bsc_sum) <- colnames(Boston)                  # Set row names
Bsc_sum
```


Catergorical criminal rate
```{r}
B_scaled$c_crim[B_scaled$crim <= -0.41] <- 1
B_scaled$c_crim[B_scaled$crim > -0.41 & B_scaled$crim <= 0] <- 2
B_scaled$c_crim[B_scaled$crim > 0 & B_scaled$crim <= 0.01] <- 3
B_scaled$c_crim[B_scaled$crim > 0.01] <- 4
B_scaled$crim <- NULL
```

Divide the data to train and test set:
```{r}
library(dplyr)
B_train <- B_scaled %>% sample_frac(.8)
names <- row.names(B_train)
B_test <- B_scaled[!row.names(B_scaled) %in% names,]
```

LDA
```{r}
m1 <- lda(c_crim ~., data = B_train) #LDA witc clustered crime rate as target
summary(m1)
```
#### 5.

The LDA plot didn't work as expected, no time to fix things:
```{r}
plot(m1, fac=m1$fac)
```
Tested the function from the course page nonetheless:
```{r}
library(RColorBrewer)
col <- brewer.pal(8, 'Set1')
palette(col)

lda.arrows <- function(x, myscale = 1, arrow_heads = 0.2, color = "orange", tex = 2, choices = c(1,2), lwd= 2){
  heads <- coef(x)
  arrows(x0 = 0, y0 = 0, 
         x1 = myscale * heads[,choices[1]], 
         y1 = myscale * heads[,choices[2]], col=color, length = arrow_heads, lwd=2)
  text(myscale * heads[,choices], labels = row.names(heads), 
       cex = tex, col=color, pos=3, lwd=2.5)
}
par(cex.label=2, cex.lab=2, cex =1.5, mar=c(5,6,3,3))
plot(m1, col=as.numeric(B_train$c_crim),  cex =1.5)
lda.arrows(m1)
```
#### 6.
Removing the classes from the test set and predicting the classes (didn't finish)
```{r}
B_test_class <- B_train
B_test$c_crim <- NULL

new_class <- predict(m1, B_test)

```




#### 7.

```{r}
Boston_scaled <- scale(Boston)

library(stats)
K4 <- kmeans(Boston_scaled, 4, iter.max = 10, nstart = 1)

```

